---
title: "2024 Daiichi Dxd"
subtitle: "Report: 30342"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
## load required packages
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
library(paletteer)
library(forcats)
```

<br>

Patient: 30342  

* Suggested Cell Types     



```{r}
dir= "~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD/"
obj.srt = readRDS(paste0(dir,"rds/P30342.24.05.07.rds"))
```


# Annotation in UMAP {.tabset}

Annotation by known cell type markers  

```{r}
orig.ident_cols = scale_color_paletteer_d("ggsci::nrc_npg") 
orig.ident_fils = scale_fill_paletteer_d("ggsci::nrc_npg") 
```

## Resolution 0.2 UMAP

```{r}
# Resolution and color palette
i=0.2
res= paste0("RNA_snn_res.", i)
palette <- wesanderson::wes_palette("FantasticFox1", length(levels(obj.srt@meta.data[,res])), type = "continuous")
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
DimPlot(obj.srt, group.by = res, cols = palette, alpha = 0.8) + 
  theme(plot.title = element_blank()) + xlab("UMAP1") + ylab("UMAP2")
```

## Resolution 0.2 UMAP (label)
```{r,fig.width=7, fig.height=6, out.width='80%'}
DimPlot(obj.srt, group.by = res, alpha = 0.8,
        label = T, label.box = T, label.size = 3, 
        cols = palette) + theme(plot.title = element_blank())
```


## Suggested cell types  

```{r}
df = obj.srt@meta.data

df = df %>% mutate(SuggestedCellType = case_when(RNA_snn_res.0.2 == 0 ~"Tcell_naive",
                                                 RNA_snn_res.0.2 == 1 ~"Macrophage1",
                                                 RNA_snn_res.0.2 == 2 ~"NK_NKT",
                                                 RNA_snn_res.0.2 == 3 ~"Bcell",
                                                 RNA_snn_res.0.2 == 4 ~"Tcell_memory_effector",
                                                 RNA_snn_res.0.2 == 5 ~"CancerCell1",
                                                 RNA_snn_res.0.2 == 6 ~"EndothelialCell",
                                                 RNA_snn_res.0.2 == 7 ~"DC",
                                                 RNA_snn_res.0.2 == 8 ~"Basophil",
                                                 RNA_snn_res.0.2 == 9 ~"Macrophage2",
                                                 RNA_snn_res.0.2 == 10 ~"Platelet",
                                                 RNA_snn_res.0.2 == 11 ~"ErythroidCell",
                                                 RNA_snn_res.0.2 == 12 ~"CancerCell2",
                                                 RNA_snn_res.0.2 == 13 ~"Neutrophil"))
```


```{r,fig.width=8.5, fig.height=6, out.width='100%'}
obj.srt[["SuggestedCellType"]] = df$SuggestedCellType
level_order = c("Tcell_naive","Macrophage1","NK_NKT","Bcell","Tcell_memory_effector",
                "CancerCell1","EndothelialCell","DC","Basophil","Macrophage2",
                "Platelet","ErythroidCell","CancerCell2","Neutrophil")
obj.srt@meta.data$SuggestedCellType = factor(obj.srt@meta.data$SuggestedCellType,
                                             levels = level_order)
DimPlot(obj.srt, group.by = "SuggestedCellType", alpha = 0.8,
        label = T, label.box = F, label.size = 4, 
        cols = palette) + theme(plot.title = element_blank())
```
<br><br>

```{r}
df %>% dplyr::select(RNA_snn_res.0.2, SuggestedCellType) %>% table() %>% data.frame() %>% filter(Freq !=0) %>% arrange(RNA_snn_res.0.2) %>% dplyr::select(RNA_snn_res.0.2, SuggestedCellType) %>% 
  DT::datatable(rownames = F, options = list(pageLength=length(levels(obj.srt@meta.data$SuggestedCellType))))
```


## Cell numbers 
```{r,fig.width=8, fig.height=5}
res = "SuggestedCellType"
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=!!sym(res))) + 
  geom_bar(alpha=0.7, color="grey5", size=0.1) +
  geom_text(stat="count", aes(label= ..count..), vjust=-0.5, size=3) +
  scale_fill_manual(values=palette) + 
  xlab("") + 
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1)) +
  ggtitle("SuggestedCellType")
```


```{r}
obj.srt@meta.data %>% dplyr::select(SuggestedCellType, orig.ident) %>% 
  table() %>% data.frame() %>% 
  tidyr::spread(orig.ident, Freq) %>% 
  DT::datatable(options = list(pageLength= length(unique(obj.srt@meta.data$SuggestedCellType))))
```


## Known Marker gene expression (Vertical Dotplot)  
```{r, fig.width=7, fig.height=12}
genes <- c("CD79A", "PTPRC", "CD38", "VPREB3", "TCL1A", "CD4", "CD8A", "CD8B","CD3D", "CD3E", "CD3G", "SELL", 
           "CD27", "IL7RA", "CCR7", "CTLA4", "IL2RA", "GZMB", "CD56", "NKp46", "CD161", "NKG7", "GZMB", "GZMA", 
           "GNLY", "LYZ","MPO","CXCR2","CD14", "FCGR3A", "CD11b", "C1QC", "S100A8", "S100A9", "CD68", "CD163", "CD80", "CD86", 
           "MRC1", "PECAM1", "CDH5", "VCAM1", "VWF", "ICAM1", "ITGAX", "CD83", "NRP1", "CLEC4C", "CD86", 
           "CD86", "ITGAX", "CD40", "CD44", "EPCAM", "KRT18")
genes <- unique(genes)

genes = genes[genes %in% rownames(obj.srt)]

my_palette <- c(colorRampPalette(colors = c("#2874A6","white"))(25), colorRampPalette(colors = c("white","red"))(30))

DotPlot(obj.srt, features = rev(genes), 
        group.by = "SuggestedCellType") +coord_flip() +
  scale_color_gradientn(colors = my_palette) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +ylab("") 
```

## Known Marker gene expression (Horizontal Dotplot)   

```{r, fig.width=12, fig.height=5, out.width="150%"}
obj.srt@meta.data$rev_suggestedCT = obj.srt@meta.data$SuggestedCellType
obj.srt@meta.data$rev_suggestedCT = factor(obj.srt@meta.data$rev_suggestedCT,
                                           levels = rev(levels(obj.srt@meta.data$SuggestedCellType)))
DotPlot(obj.srt, features = genes, 
        group.by = "rev_suggestedCT") +
  scale_color_gradientn(colors = my_palette) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +ylab("") 
```


## Reference    

| Gene    | Cell Type                             |
|---------|---------------------------------------|
| CD79A   | B cells                               |
| PTPRC   | All leukocytes (B cells, T cells, etc.) |
| CD38    | Activated B cells, plasma cells       |
| VPREB3  | B cell development                    |
| TCL1A   | B cells (associated with B cell lymphoma) |
| CD4     | Helper T cells                        |
| CD8A    | Cytotoxic T cells                     |
| CD8B    | Cytotoxic T cells                     |
| CD3D    | All T cells                           |
| CD3E    | All T cells                           |
| CD3G    | All T cells                           |
| SELL    | Naive T cells                         |
| CD27    | Memory B cells, plasma cells, T cells |
| CCR7    | Naive and central memory T cells      |
| CTLA4   | Activated T cells                     |
| IL2RA   | Activated T cells, regulatory T cells |
| GZMB    | Cytotoxic T cells, NK cells           |
| NKG7    | Cytotoxic T cells, NK cells           |
| GZMA    | Cytotoxic T cells, NK cells           |
| GNLY    | NK cells, cytotoxic T cells           |
| LYZ     | Macrophages, monocytes, other leukocytes |
| MPO     | Neutrophils                           |
| CXCR2   | Neutrophils, some monocytes           |
| CD14    | Monocytes, macrophages                |
| FCGR3A  | NK cells, monocytes, macrophages      |
| C1QC    | Monocytes, macrophages                |
| S100A8  | Inflammatory monocytes, macrophages   |
| S100A9  | Inflammatory monocytes, macrophages   |
| CD68    | Macrophages, monocytes                |
| CD163   | Macrophages, monocytes                |
| CD80    | Dendritic cells, B cells              |
| CD86    | Dendritic cells, B cells              |
| MRC1    | Macrophages, dendritic cells          |
| PECAM1  | Endothelial cells, vascular cells     |
| CDH5    | Vascular endothelial cells            |
| VCAM1   | Vascular endothelial cells            |
| VWF     | Vascular endothelial cells, platelets |
| ICAM1   | Various cell types (involved in inflammation) |
| ITGAX   | Dendritic cells, macrophages          |
| CD83    | Mature dendritic cells                |
| NRP1    | Dendritic cells, some immune cells    |
| CLEC4C  | Dendritic cells                       |
| CD40    | B cells, macrophages, dendritic cells |
| CD44    | Cancer cells                          |
| EPCAM   | Cancer cells                          |
| KRT18   | Cancer cells                          |

<br><br><br>

# Manual annotation by Cluster Markers {.tabset} 


## Table  

**Cell Types and MK Genes by Cluster**   

| Cluster | Inferred Cell Type               | Key Genes                                                      |
|---------|----------------------------------|---------------------------------------------------------------|
| 0       | T cells                          | CD3D, CD2, GIMAP7, CD3G, CD3E, TRAC, BCL11B, ...               |
| 1       | Macrophages 1                    | MS4A7, C5AR1, PLA2G7, SPI1, PILRA, KYNU, DMXL2, ...            |
| 2       | NK cells                         | NKG7, GNLY, KLRD1, GZMB, CST7, TYROBP, APOBEC3G, ...           |
| 3       | B cells                          | MS4A1, CD79A, BANK1, TNFRSF13C, HLA-DQB1, CD74, ...            |
| 4       | Regulatory T cells (Tregs)       | TIGIT, IL2RA, LINC01943, TNFRSF4, FOXP3, CTLA4, ...            |
| 5       | Epithelial cells (Cancer) 1      | ELF3, EPCAM, MAL2, PAWR, WFDC2, MYO1B, KRT18, ...              |
| 6       | Epithelial cells (Cancer) 2      | TM4SF1, KRT18, S100A13, CCND1, IGFBP7, S100A10, ...            |
| 7       | Dendritic cells                  | LAMP3, GPR157, IDO1, CSF2RA, RAMP1, CD86, DAP12, ...           |
| 8       | Mast cells                       | TPSAB1, TPSB2, GATA2, HPGDS, CPA3, PRG2, KIT, ...              |
| 9       | Macrophages 2                    | MARCO, VSIG4, FABP4, C1QC, PLBD1, TREM1, MSR1, ...             |
| 10      | Fibroblasts                      | COL6A2, IGFBP7, FGF2, C11orf96, FERMT2, CALD1, ...             |
| 11      | Red blood cells                  | HBA2, HBA1, HBB, SLC25A37                                       |
| 12      | Alveolar epithelial cells        | SFTPB, DDAH1, SFTPD, SFTPA1, SDR16C5, SFTPA2, ...              |
| 13      | Neutrophils                      | PI3, FCGR3B, AL021978.1, HCAR3, KCNJ15, S100P, ...             |


## Cell number   


```{r}
inferred_cell_types <- c(
  "T cells",                       # Cluster 0
  "Macrophages1",                   # Cluster 1
  "NK cells",                      # Cluster 2
  "B cells",                       # Cluster 3
  "Regulatory T cells (Tregs)",    # Cluster 4
  "Epithelial cells (Cancer) 1",              # Cluster 5
  "Epithelial cells (Cancer) 2",              # Cluster 6
  "Dendritic cells",               # Cluster 7
  "Mast cells",                    # Cluster 8
  "Macrophages2",                   # Cluster 9
  "Fibroblasts",                   # Cluster 10
  "Red blood cells",               # Cluster 11
  "Alveolar epithelial cells",     # Cluster 12
  "Neutrophils"                    # Cluster 13
)

meta_data <- obj.srt@meta.data

meta_data <- meta_data %>%
  mutate(
    Manual_cell_type = ifelse(RNA_snn_res.0.2 == 0, inferred_cell_types[1],
                                ifelse(RNA_snn_res.0.2 == 1, inferred_cell_types[2],
                                       ifelse(RNA_snn_res.0.2 == 2, inferred_cell_types[3],
                                              ifelse(RNA_snn_res.0.2 == 3, inferred_cell_types[4],
                                                     ifelse(RNA_snn_res.0.2 == 4, inferred_cell_types[5],
                                                            ifelse(RNA_snn_res.0.2 == 5, inferred_cell_types[6],
                                                                   ifelse(RNA_snn_res.0.2 == 6, inferred_cell_types[7],
                                                                          ifelse(RNA_snn_res.0.2 == 7, inferred_cell_types[8],
                                                                                 ifelse(RNA_snn_res.0.2 == 8, inferred_cell_types[9],
                                                                                        ifelse(RNA_snn_res.0.2 == 9, inferred_cell_types[10],
                                                                                               ifelse(RNA_snn_res.0.2 == 10, inferred_cell_types[11],
                                                                                                      ifelse(RNA_snn_res.0.2 == 11, inferred_cell_types[12],
                                                                                                             ifelse(RNA_snn_res.0.2 == 12, inferred_cell_types[13],
                                                                                                                    ifelse(RNA_snn_res.0.2 == 13, inferred_cell_types[14], NA)))))))))))))))


meta_data$Manual_cell_type = factor(meta_data$Manual_cell_type, 
                                      levels = inferred_cell_types)
obj.srt@meta.data <- meta_data 
```

```{r}
obj.srt@meta.data %>% dplyr::select(Manual_cell_type, orig.ident) %>% 
  table() %>% data.frame() %>% 
  tidyr::spread(orig.ident, Freq) %>% 
  DT::datatable(options = list(pageLength= length(unique(obj.srt@meta.data$Manual_cell_type))))
```


```{r}
# List of top genes for each cluster
top_genes <- c('CD3D', 'CD2', 'GIMAP7', 'CD3G', 'CD3E', 
               'MS4A7', 'C5AR1', 'PLA2G7', 'SPI1', 'PILRA', 
               'NKG7', 'GNLY', 'KLRD1', 'GZMB', 'CST7', 
               'MS4A1', 'CD79A', 'BANK1', 'TNFRSF13C', 'HLA-DQB1', 
               'TIGIT', 'IL2RA', 'LINC01943', 'TNFRSF4', 'FOXP3', 
               'ELF3', 'EPCAM', 'MAL2', 'PAWR', 'WFDC2', 
               'TM4SF1', 'KRT18', 'S100A13', 'CCND1', 'IGFBP7', 
               'LAMP3', 'GPR157', 'IDO1', 'CSF2RA', 'RAMP1', 
               'TPSAB1', 'TPSB2', 'GATA2', 'HPGDS', 'CPA3', 
               'MARCO', 'VSIG4', 'FABP4', 'C1QC', 'PLBD1', 
               'COL6A2', 'IGFBP7', 'FGF2', 'C11orf96', 'FERMT2', 
               'HBA2', 'HBA1', 'HBB', 'SLC25A37', 
               'SFTPB', 'DDAH1', 'SFTPD', 'SFTPA1', 'SDR16C5', 
               'PI3', 'FCGR3B', 'AL021978.1', 'HCAR3', 'KCNJ15')

top_genes <- top_genes[!duplicated(top_genes)]
# Plot DotPlot
# DotPlot(obj.srt, features = top_genes, group.by = "RNA_snn_res.0.2") + coord_flip()
```

## Dotplot (Vertical)     

```{r, fig.width=8, fig.height=16, eval=FALSE}
my_palette <- c(colorRampPalette(colors = c("#2874A6","white"))(12), colorRampPalette(colors = c("white","red"))(30))

DotPlot(obj.srt, features = rev(top_genes), 
        group.by = "RNA_snn_res.0.2") +coord_flip() +
  scale_color_gradientn(colors = my_palette) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 0, hjust = 1)) +
  xlab("") +ylab("") 
```



```{r, fig.width=8, fig.height=16}
my_palette <- c(colorRampPalette(colors = c("#2874A6","white"))(12), colorRampPalette(colors = c("white","red"))(30))

DotPlot(obj.srt, features = rev(top_genes), 
        group.by = "Manual_cell_type") +coord_flip() +
  scale_color_gradientn(colors = my_palette) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +ylab("") 
```


## Dotplot (Horizontal)     

```{r, fig.width=16, fig.height=8}

obj.srt$inferred_cell_type = factor(obj.srt$Manual_cell_type, 
                                      levels = rev(inferred_cell_types))
my_palette <- c(colorRampPalette(colors = c("#2874A6","white"))(12), colorRampPalette(colors = c("white","red"))(30))

DotPlot(obj.srt, features = top_genes, 
        group.by = "Manual_cell_type") +
  scale_color_gradientn(colors = my_palette) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +ylab("") 
```



<!-- **Macrophage Marker Genes and Their Roles**   -->

<!-- | Gene    | Role                                                                                      | Macrophage Annotation                                      | -->
<!-- |---------|-------------------------------------------------------------------------------------------|-----------------------------------------------------------| -->
<!-- | MS4A7   | Membrane protein involved in macrophage differentiation and function.                     | High expression in mature macrophages.                    | -->
<!-- | C5AR1   | Receptor in the complement system that regulates inflammatory responses and chemotaxis.   | Indicates involvement in inflammatory responses.          | -->
<!-- | PLA2G7  | Enzyme that produces pro-inflammatory mediators, involved in lipid metabolism.            | Reflects activity in inflammation and lipid metabolism.   | -->
<!-- | SPI1    | Transcription factor essential for macrophage differentiation and maturation.             | Critical role in macrophage development and function.     | -->
<!-- | PILRA   | Immune receptor involved in cell signaling and regulation of immune responses.            | Regulates macrophage immune responses.                    | -->
<!-- | KYNU    | Enzyme in the tryptophan metabolism pathway, involved in immune regulation and inflammation. | Reflects metabolic activity and role in inflammatory responses. | -->
<!-- | DMXL2   | Protein associated with various cellular functions, including signaling and metabolism.   | Involved in diverse metabolic processes and cell functions. | -->


<br><br><br>