---
title: "Brain trauma study"
subtitle: "works.io"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br><hr>

# Papaer

* Single-nucleus profiling of adult mice sub-ventricular zone after blast-related traumatic brain injury

## 7 SVZ cell types

### 1. Neurons
  * Meg3
  * Syt1, Snap25
  
### 2. NSC-astrocytes (Neuronal Stem Cells)
  * Nr2e1, Hes5, Slc1a2

### 3. Oligodendrocytes
  * Mog, Aspa

### 4. OPCs (progenitor cells)
  * Pdgfra 
  
### Microglia
  * Csf1r, Tmem119, P2ry12
  
### Ependymal Cells
  * Dnah11, Spag16, Admts20
  
### Endothelium-mural cells
  * Pecam1, Vwf, Vtn

```{r}
cell_data <- data.frame(
  Cell_type = c("Neuron", "Microglia", "NSC-Astrocyte", "Oligodendrocyte", "OPC", 
                "Ependymal", "Endothelial-Mural"),
  Markers = c(
    "Meg3, Syt1, Snap25",
    "Csf1r, Tmem119, P2ry12",
    "Slc1a3, Apoe, Aldoc, Aqp4",
    "Mog, Aspa, Tubb4a, Apod",
    "Pdgfra",
    "Dnah11, Spag16, Adamts20, Ak7, Ak9, Armc3",
    "Pecam1(CD31), Vwf, Vtn"
  ),
  stringsAsFactors = FALSE # 문자형 데이터 처리를 위해 문자열로 유지
)

print(cell_data)
```


<hr>

## DEG cutoff 
  * lnFC > 0.25, P < 0.05, Min. PCT > 0.1


<hr>
<br><br><br><br>
<br><br><br><br>